A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-04-23 paper.Rmd
done with 5.2

The power comparison demonstrated that the appropriate conventional tests will reject more aggressively than visual tests, but we don't know how the decisions for each lineup would agree or disagree. Here we compare the reject or fail to reject decisions of these tests, across all the lineups. Figure \ref{fig:p-value-comparison} shows the agreement of the conventional and visual tests using a mosaic plot for both non-linearity patterns and heteroskedasticity patterns.

For both patterns the lineups resulting in a rejection by the visual test are *all* also rejected by the conventional test, except for one from the heteroskedasticity model. This reflects exactly the story from the previous section, that the conventional tests reject more aggressively than the visual test.

For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time. Of the lineups rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test. There are no lineups that are rejected by the visual test but not by the conventional test.

In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests. The visual test rejects `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the lineups that the conventional visual test reject.

@@

Surprisingly, the visual test rejects `r nr$n[1]` of the `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject. Figure \ref{fig:heter-example} shows this lineup. The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e(E)=4.02$). This is reflected by the visual test $p\text{-value} = 0.026$. But the BP test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$. This lineup was evaluated by 11 participants, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values. It may have been the small sample size and the presence of a few outliers that may have resulted in the lack of detection by the conventional test.

@@

Because the power curve of the visual tests are a shift to the right of the conventional test (Figures \ref{fig:polypower} and \ref{fig:heter-example}) it would be curious to know whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test. Interestingly, it doesn't. Despite resulting less rejections, neither the RESET or BP tests come to complete agreement with the visual test (See Appendix).

>>>>

The power comparison demonstrates that the appropriate conventional tests will reject more aggressively than visual tests, but we don't know how the decisions for each lineup would agree or disagree. Here we compare the reject or fail to reject decisions of these tests, across all the lineups. Figure \ref{fig:p-value-comparison} shows the agreement of the conventional and visual tests using a mosaic plot for both non-linearity patterns and heteroskedasticity patterns. For both patterns the lineups resulting in a rejection by the visual test are *all* also rejected by the conventional test, except for one from the heteroskedasticity model. This reflects exactly the story from the previous section, that the conventional tests reject more aggressively than the visual test.

For non-linearity lineups, conventional tests and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time, respectively. Of the lineups rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test. There are no lineups that are rejected by the visual test but not by the conventional test.

In heteroskedasticity lineups, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests. Of the lineups rejected by the conventional test, the visual test rejects more than two-thirds of them, too.

@@

Surprisingly, the visual test rejects `r nr$n[1]` of the `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject. Figure \ref{fig:heter-example} shows this lineup. The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e(E)=4.02$), which is reflected by the visual test $p\text{-value} = 0.026$. But the BP test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$. This lineup was evaluated by 11 participants, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values. It may have been the small sample size and the presence of a few outliers that may have resulted in the lack of detection by the conventional test.

@@

Because the power curve of the visual tests are a shift to the right of the conventional test (Figures \ref{fig:polypower} and \ref{fig:heter-example}) we examined whether adjusting the significance level (to .001, .0001, .00001, ...) of the conventional test would generate similar decisions to that of the visual test. Interestingly, it doesn't: despite resulting less rejections, neither the RESET or BP tests come to complete agreement with the visual test (see Appendix).

### 2023-04-23 paper.tex
done with 5.2

The power comparison demonstrated that the appropriate conventional
tests will reject more aggressively than visual tests, but we don't know
how the decisions for each lineup would agree or disagree. Here we
compare the reject or fail to reject decisions of these tests, across
all the lineups. Figure \ref{fig:p-value-comparison} shows the agreement
of the conventional and visual tests using a mosaic plot for both
non-linearity patterns and heteroskedasticity patterns.

For both patterns the lineups resulting in a rejection by the visual
test are \emph{all} also rejected by the conventional test, except for
one from the heteroskedasticity model. This reflects exactly the story
from the previous section, that the conventional tests reject more
aggressively than the visual test.

For lineups containing non-linearity patterns, conventional tests reject
69\% and visual tests reject 32\% of the time. Of the lineups rejected
by the conventional test, 46\% are rejected by the visual test, that is,
approximately half as many as the conventional test. There are no
lineups that are rejected by the visual test but not by the conventional
test.

In terms of lineups containing heteroskedasticity patterns, 76\% are
rejected by conventional tests, while 56\% are rejected by visual tests.
The visual test rejects 73\% of the lineups that the conventional visual
test reject.

Surprisingly, the visual test rejects 1 of the 33 (3\%) of lineups where
the conventional test does not reject. Figure \ref{fig:heter-example}
shows this lineup. The data plot in position seventeen displays a
relatively strong heteroskedasticity pattern, and has a strong effect
size (\(log_e(E)=4.02\)). This is reflected by the visual test
\(p\text{-value} = 0.026\). But the BP test \(p\text{-value} = 0.056\),
is slightly above the significance cutoff of \(0.05\). This lineup was
evaluated by 11 participants, it has experimental factors \(a = 0\)
(``butterfly'' shape), \(b = 64\) (large variance ratio), \(n = 50\)
(small sample size), and a uniform distribution for the fitted values.
It may have been the small sample size and the presence of a few
outliers that may have resulted in the lack of detection by the
conventional test.

@@

Because the power curve of the visual tests are a shift to the right of
the conventional test (Figures \ref{fig:polypower} and
\ref{fig:heter-example}) it would be curious to know whether adjusting
the significance level (to .001, .0001, .00001, \ldots) of the
conventional test would generate similar decisions to that of the visual
test. Interestingly, it doesn't. Despite resulting less rejections,
neither the RESET or BP tests come to complete agreement with the visual
test (See Appendix).

>>>>

The power comparison demonstrates that the appropriate conventional
tests will reject more aggressively than visual tests, but we don't know
how the decisions for each lineup would agree or disagree. Here we
compare the reject or fail to reject decisions of these tests, across
all the lineups. Figure \ref{fig:p-value-comparison} shows the agreement
of the conventional and visual tests using a mosaic plot for both
non-linearity patterns and heteroskedasticity patterns. For both
patterns the lineups resulting in a rejection by the visual test are
\emph{all} also rejected by the conventional test, except for one from
the heteroskedasticity model. This reflects exactly the story from the
previous section, that the conventional tests reject more aggressively
than the visual test.

For non-linearity lineups, conventional tests and visual tests reject
69\% and 32\% of the time, respectively. Of the lineups rejected by the
conventional test, 46\% are rejected by the visual test, that is,
approximately half as many as the conventional test. There are no
lineups that are rejected by the visual test but not by the conventional
test.

In heteroskedasticity lineups, 76\% are rejected by conventional tests,
while 56\% are rejected by visual tests. Of the lineups rejected by the
conventional test, the visual test rejects more than two-thirds of them,
too.

Surprisingly, the visual test rejects 1 of the 33 (3\%) of lineups where
the conventional test does not reject. Figure \ref{fig:heter-example}
shows this lineup. The data plot in position seventeen displays a
relatively strong heteroskedasticity pattern, and has a strong effect
size (\(log_e(E)=4.02\)), which is reflected by the visual test
\(p\text{-value} = 0.026\). But the BP test \(p\text{-value} = 0.056\),
is slightly above the significance cutoff of \(0.05\). This lineup was
evaluated by 11 participants, it has experimental factors \(a = 0\)
(``butterfly'' shape), \(b = 64\) (large variance ratio), \(n = 50\)
(small sample size), and a uniform distribution for the fitted values.
It may have been the small sample size and the presence of a few
outliers that may have resulted in the lack of detection by the
conventional test.

@@

Because the power curve of the visual tests are a shift to the right of
the conventional test (Figures \ref{fig:polypower} and
\ref{fig:heter-example}) we examined whether adjusting the significance
level (to .001, .0001, .00001, \ldots) of the conventional test would
generate similar decisions to that of the visual test. Interestingly, it
doesn't: despite resulting less rejections, neither the RESET or BP
tests come to complete agreement with the visual test (See Appendix).
